### 2014-11-11 tex/bisimulation.tex
Fixed a few typos

Let $p \rgse p^\prime$ be defined as a union over state relations:
$$
\rgse = \rgse^\mathit{A} \cup  \rgse^\mathit{A\cdot} \cup \rgse^\mathit{W} \cup \rgse^\mathit{W\cdot} \cup \rgse^\mathit{E} \cup \rgse^\mathit{E\cdot} \cup \rgse^\com
$$
where \emph{A} is a field access $\lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f}
\rightarrow \cfgnt{k}\rp$, \emph{W} is a field write $\lp
\cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgnt{*} \rightarrow
\cfgnt{k}\rp$, \emph{E} is a reference compare $\lp
\cfgnt{v}\ \cfgt{=}\ \cfgnt{*} \rightarrow \cfgnt{k}\rp$, $\cdot$
indicates a null reference in the operation or a false outcome, and
\emph{\com} is everything else in the language. Any state relation, say $\rightarrow_x$, is trivially extended to sets of states as
$$
P \hookrightarrow_x P^\prime \Longleftrightarrow \forall p \in P\ (\forall p^\prime\ (p \rightarrow_x p^\prime \Rightarrow p^\prime \in P^\prime))
$$
Let $\rgse$ be the extension of $\rgse$ to sets of states.
$$
\rsgse = \hookrightarrow_\gse^\mathit{A}
\cup \hookrightarrow_\gse^\mathit{A\cdot} \cup \hookrightarrow_\gse^\mathit{W} \cup
\hookrightarrow_\gse^\mathit{W\cdot} \cup \hookrightarrow_\gse^\mathit{E} \cup \hookrightarrow_\gse^\mathit{E\cdot}
\cup \hookrightarrow_\gse^\com
$$
The definition embodies the notion of splitting equivalence classes at
certain operations. The set $P$ at a field access for example has two
potential successors: a non-null and null reference. Respectively, the
set $P \rgse^\com$ only has a single successor since those expressions
are not related to control flow.

@@

\begin{definition}
\label{def:hc}
The \textbf{homomorphism constraint} is
\begin{align*}
\mathbb{HC}(p \rightharpoonup_{h} q) &= \\
 \bigwedge \{ \phi_b\ | \exists (\phi_a\ l) \in \cfgnt{L}_p^\rightarrow ( (\phi_b\ h(l)) \in \cfgnt{L}_q^\rightarrow)\} 
\end{align*}
\end{definition}
The functional equivalence between heaps on a homomorphism and
constraint asserts a common structures in the two heaps under certain
conditions. The equivalence is used to related states in $p \rgse
p^\prime$ to states in $q \rsym q^\prime$.
\begin{definition}
\label{representation}
States $(p\ q)$ are in the \textbf{representation relation}, $p \sqsubset q$, if and only if, $\eta_p = \eta_q ,\ \cfgnt{e}_p =
\cfgnt{e}_q ,\ \cfgnt{k}_p = \cfgnt{k}_q$, and there exists a
homomorphism $p \rightharpoonup_{h}
q$ such that
\begin{equation}
\label{eqn:valid}
 \mathbb{S}( \phi_q \wedge \mathbb{HC}(s_p \rightharpoonup_{h} s_q) ) 
\end{equation}
The represented relation is extended to sets of states $P$ and a single state $q$ as\footnote{The full proof argues $P \sqsubset Q$ but that $Q$ is
  always a singleton set containing $q$.}
$$
P \sqsubset q \Longleftrightarrow \forall p\ (p \sqsubset q \Rightarrow p \in P)
$$
\end{definition}
The statement $p \sqsubset q$ ensures that a functionally equivalent
heap to the one in $p$ is present, by the homomorphism, and valid, by
the heap constraint and path constraint, in $q$. As the states in $P$
are only differentiated by heaps and those states are only
differentiated from $q$ by both the heap and path constraint in $q$,
$P \sqsubset q$ implies that $q$ is representative of all the states
in $P$ up to the given point of execution expressed in $\phi_q$.
\begin{definition}
\label{bisimulation}
The \textbf{functional associated to bisimulation} applied to $\sqsubset$, denoted as $F_\sim(\sqsubset)$, is the set of all pairs
$(P\ q)$ such that
\begin{equation}
\label{eqn:BisimulationForwards}
\forall P^\prime\ ( P \rsgse P^\prime \Rightarrow \exists q^\prime\ ( (q \rsym q^\prime )\wedge (P^\prime\ \sqsubset\ q^\prime )))
\end{equation}
\begin{equation}
\label{eqn:BisimulationBackwards}
\forall q^\prime\ ( q \rsym q^\prime\Rightarrow \exists P^\prime\ ( (P \rsgse P^\prime )\wedge (P^\prime\ \sqsubset\ q^\prime )))
\end{equation}
If $\sqsubset$ is a bisimulation, then the greatest fixed point of $F_\sim(\sqsubset)$ is the bisimilarity relation.
\end{definition}
As is typical, the functional reasons over a forward and backward
simulation, while the application to dissimilar states using a
meta-relation, $\rsgse$, is somewhat non-standard; though convenient
in this instance since it captures the notion of $q$ as a set of
states \cite{GSE:barbedbisimulation}.

@@

\begin{theorem}
\label{th:bisim}
The relation $\sqsubset$ is a bisimulation.
\end{theorem}
\begin{proof}
By definition of the bisimilarity relation. All of the common rules
make no changes to the heap, so $P \subset q$ is included in the
functional. The other transition relations are supported by lemmas
such as \lemref{lem:access}.
\end{proof}

>>>>

Let $p \rgse p^\prime$ be defined as a union over state relations:
$$
\rgse = \rgse^\mathit{A} \cup  \rgse^\mathit{A^\prime} \cup \rgse^\mathit{W} \cup \rgse^\mathit{W^\prime} \cup \rgse^\mathit{E} \cup \rgse^\mathit{E\cdot} \cup \rgse^\com
$$
where \emph{A} is a field access $\lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f}
\rightarrow \cfgnt{k}\rp$, \emph{W} is a field write $\lp
\cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgnt{*} \rightarrow
\cfgnt{k}\rp$, \emph{E} is a reference compare $\lp
\cfgnt{v}\ \cfgt{=}\ \cfgnt{*} \rightarrow \cfgnt{k}\rp$, $\cdot$
indicates a null reference in the operation or a false outcome, and
\emph{\com} is everything else in the language. Any state relation, say $\rightarrow_x$, is trivially extended to sets of states as
$$
P \hookrightarrow_x P^\prime \Longleftrightarrow \forall p \in P\ (\forall p^\prime\ (p \rightarrow_x p^\prime \Rightarrow p^\prime \in P^\prime))
$$
Let $\hookrightarrow_\gse$ be the extension of $\rgse$ to sets of states.
$$
\rsgse = \hookrightarrow_\gse^\mathit{A}
\cup \hookrightarrow_\gse^\mathit{A\cdot} \cup \hookrightarrow_\gse^\mathit{W} \cup
\hookrightarrow_\gse^\mathit{W\cdot} \cup \hookrightarrow_\gse^\mathit{E} \cup \hookrightarrow_\gse^\mathit{E\cdot}
\cup \hookrightarrow_\gse^\com
$$
The definition embodies the notion of splitting equivalence classes at
certain operations. The set $P$ at a field access for example has two
potential successors: a non-null and null reference. Respectively, the
set $P \rgse^\com P^\prime$ only has a single successor since those expressions
are not related to control flow.

@@

\begin{definition}
\label{def:hc}
The \textbf{homomorphism constraint} is
\begin{align*}
\mathbb{HC}(p \rightharpoonup_{h} q) &= \\
 \bigwedge \{ \phi_b\ | \exists (\phi_a\ l) \in \cfgnt{L}_p^\rightarrow ( (\phi_b\ h(l)) \in \cfgnt{L}_q^\rightarrow)\} 
\end{align*}
\end{definition}
The functional equivalence between heaps on a homomorphism and
a homomorphism constraint asserts a common structures in the two heaps under certain
conditions. The equivalence is used to related states in $p \rgse
p^\prime$ to states in $q \rsym q^\prime$.
\begin{definition}
\label{representation}
States $(p\ q)$ are in the \textbf{representation relation}, $p \sqsubset q$, if and only if, $\eta_p = \eta_q ,\ \cfgnt{e}_p =
\cfgnt{e}_q ,\ \cfgnt{k}_p = \cfgnt{k}_q$, and there exists a
homomorphism $p \rightharpoonup_{h}
q$ such that
\begin{equation}
\label{eqn:valid}
 \mathbb{S}( \phi_q \wedge \mathbb{HC}(s_p \rightharpoonup_{h} s_q) ) 
\end{equation}
The represented relation is extended to sets of states $P$ and a single state $q$ as\footnote{The full proof argues $P \sqsubset Q$ but that $Q$ is
  always a singleton set containing $q$.}
$$
P \sqsubset q \Longleftrightarrow \forall p\ (p \sqsubset q \Rightarrow p \in P)
$$
\end{definition}
The statement $p \sqsubset q$ ensures that a functionally equivalent
heap to the one in $p$ is present, by the homomorphism, and valid, by
the heap constraint and path constraint, in $q$. As the states in $P$
are only differentiated by heaps and those states are only
differentiated from $q$ by both the heap and path constraint in $q$,
$P \sqsubset q$ implies that $q$ is representative of all the states
in $P$ up to the given point of execution expressed in $\phi_q$.
\begin{definition}
\label{bisimulation}
The \textbf{functional associated to bisimulation} applied to $\sqsubset$, denoted as $F_\sim(\sqsubset)$, is the set of all pairs
$(P\ q)$ such that
\begin{equation}
\label{eqn:BisimulationForwards}
\forall P^\prime\ ( P \rsgse P^\prime \Rightarrow \exists q^\prime\ ( q \rsym q^\prime \wedge P^\prime\ \sqsubset\ q^\prime))
\end{equation}
\begin{equation}
\label{eqn:BisimulationBackwards}
\forall q^\prime\ ( q \rsym q^\prime\Rightarrow \exists P^\prime\ ( P \rsgse P^\prime \wedge P^\prime\ \sqsubset\ q^\prime))
\end{equation}
If $\sqsubset$ is a bisimulation, then the greatest fixed point of $F_\sim(\sqsubset)$ is the bisimilarity relation denoted by $\sim$.
\end{definition}
As is typical, the functional reasons over a forward and backward
simulation, while the application to dissimilar states using a
meta-relation, $\rsgse$, is somewhat non-standard; though convenient
in this instance since it captures the notion of $q$ as a set of
states \cite{GSE:barbedbisimulation}.

@@

\begin{theorem}
\label{th:bisim}
The relation $\sqsubset$ is a bisimulation: $\sqsubset\ \subseteq\ \sim$
\end{theorem}
\begin{proof}
By definition of the bisimilarity relation. All of the common rules
make no changes to the heap, so $P \subset q$ is included in the
functional. The other transition relations are supported by lemmas
such as \lemref{lem:access}.
\end{proof}
